Exploring the Vectorization of Python Constructs Using Pythran and Boost SIMD

### 2014-01-19 doc/papers/wpmvp14/paper.tex
Take into account reviewers comment
Most notably, emphasise the fact we are using doubles and not floats
during the experiments...

@@

    The Python language is highly dynamic, most notably due to late binding. As
    a consequence, programs run using Python typically run an order of magnitude
    slower than their C counterpart. It is also a high level language whose
    semantic can be made more static without much change from a user point of
    view in the case of mathematical applications. In that case, the language
    provides several vectorization opportunities that are studied in this
    paper, and evaluated in the context of Pythran, an ahead-of-time compiler
    that turns Python module into C++ meta-programs.

@@

A native module is a shared library whose interface make it callable from the
Python interpreter using the regular \texttt{import} mechanism. It makes it
possible to call native code from Python in a transparent way. Typically Python
object are unboxed when transfered from Python to C, where efficient
computations can happen. Then the result is boxed when getting back to the
Python world. In between, vectorization can happen. This is the approach taken
by the \texttt{numpy} module: a native type, the \texttt{ndarray} contains
contiguous unboxed elements, say single precision floating point numbers, and
provides a great deal of high-level functions to manipulate them. All the
computation-intensive part is done in pure C.  Vectorization can eventually
happen at this level, for instance when performing a point-to-point operation
like the sum of two arrays. However the granularity of the operations limits
the ratio of \texttt{load}/\texttt{store} versus operation. For instance, if
\texttt{a}, \texttt{b} and \texttt{c} are 1-D arrays, the sum \texttt{a + b +
c} is computed using two loops and an intermediate array. It means there are 4
\texttt{load}/\texttt{store} and only two additions per loop iteration and vectorization is
unlikely to be profitable.

@@

>>>>

%\doi{}

@@

    The Python language is highly dynamic, most notably due to late binding. As
    a consequence, programs using Python typically run an order of magnitude
    slower than their C counterpart. It is also a high level language whose
    semantic can be made more static without much change from a user point of
    view in the case of mathematical applications. In that case, the language
    provides several vectorization opportunities that are studied in this
    paper, and evaluated in the context of Pythran, an ahead-of-time compiler
    that turns Python module into C++ meta-programs.

@@

A native module is a shared library whose interface makes it callable from the
Python interpreter using the regular \texttt{import} mechanism. It makes it
possible to call native code from Python in a transparent way. Typically Python
objects are unboxed when transfered from Python to C, where efficient
computations can happen. Then the result is boxed when getting back to the
Python world. In between, vectorization can happen. This is the approach taken
by the \texttt{numpy} module: a native type, the \texttt{ndarray} contains
contiguous unboxed elements, say single precision floating point numbers, and
provides a great deal of high-level functions to manipulate them. All the
computation-intensive part is done in pure C.  Vectorization can eventually
happen at this level, for instance when performing a point-to-point operation
like the sum of two arrays. However the granularity of the operations limits
the ratio of \texttt{load}/\texttt{store} versus operation. For instance, if
\texttt{a}, \texttt{b} and \texttt{c} are 1-D arrays, the sum \texttt{a + b +
c} is computed using two loops and an intermediate array. It means there are 4
\texttt{load}/\texttt{store} and only two additions per loop iteration and vectorization is
unlikely to be profitable.

@@

All the benchmarks use Numpy arrays of \emph{double precision} random floats as
input, which means a maximum theoritical speedup of $\times4$ using AVX.
